Simultaneous Elicitation of Scoring Rule and Agent Preferences for Robust Winner Determination

### 2021-01-15 minimax.tex
Integrated some display math. Poor reader.

Let $V$ be the set of possible preference profiles (the union, for any integer $n$, of the $n$-fold Cartesian product of the linear orders over the alternatives).
A social choice function $f : V \rightarrow \powersetz{A}$ associates a profile with a set of winners, where $\powersetz{A}$ represents the set of subsets of $A$ except for the empty set (sets are used for tied winners).
Among the many possible social choice functions, we consider {\em positional scoring rules (PSR)}, which attach weights to positions according to the vector $(w_1, \ldots, w_m)$, also known as the {\em scoring vector}.
An alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
	\label{eq:srule}
	s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
	= \sum_{r=1}^{m} \alpha^{x}_r w_r 
\end{align}
where $\alpha^{x}_r$ is the number of times that alternative $x$ was ranked in the $r$-th position.
The winners are the alternatives with the highest score.

In this work we assume fixed, but unknown to us, a profile $\profile^*$ representing the preferences $\pref_j^*$ of the agents, and a weight vector $\w^*$, representing the preferences of the chair. We want to reason about partial preference information concerning those objects.
At a given time, our knowledge of agent $j$'s preference is encoded by a partial order over the alternatives, thus a transitive and asymmetric binary relation, denoted by $\ppref_j$. 
In this work we assume that preference information is truthful, i.e. $a \ppref_j b ⇒ a \pref_j^* b$.
%We use $\prefinc$ to denote incomparability, that is $a \prefinc_{j} b$ iff $a \nppref_j b \wedge b \nppref_j a$.
An incomplete profile $\pprofile = (\ppref_1, \ldots, \ppref_n)$ maps each agent to a partial preference.

@@

We also assume that the weights of the scoring rule are only partially specified.
Therefore, the vector $(w_1,\ldots,w_m)$  is not known but we are given a set of constraints restraining the possible values that weights can take.
We consider a decreasing sequence of weights:
\begin{align}
	1=w_{1} ≥ w_{2} ≥ \ldots ≥ w_{m}=0. \label{eq:monotone}
\end{align}
This is a natural assumption, as it is better to be ranked first than second, second than third, etc. 
Without loss of generality, we consider that $w_1=1$ and $w_m=0$.

@@

\section{Robust winner determination}
\label{sec:mmr}
In this paper, we consider a setting where both the agent preferences and the preferences of the chair about the voting rule are incomplete.
As reviewed in \cref{sec:related}, some authors have considered possible and necessary winners assuming 
either a partial profile or assuming an incompletely specified scoring rule.

Notice that often there are no necessary winners and too many possible winners; it is therefore useful to declare a winner given partial information.
As a decision criterion to determine a winner, we propose to use minimax regret. 
Minimax regret \citep{Savage1954} is a decision criterion that has been used for robust optimization under data uncertainty \citep{Kouvelis1997} as well as in decision-making with uncertain utility values \citep{Salo2001,Boutilier2006}.
In particular, \citet{Lu2011} have adopted minimax regret for winner determination in social choice where
the preferences of agents that are only partially known, while the social choice function is predetermined and known.

@@

The maximum regret is considered by assuming that an adversary can both 1) extend the partial profile $\pprofile$ into a complete profile, and 2) instantiate the weights choosing among any weight vector in $\pw$, where $\pprofile$ and $\pw$ represent our knowledge so far.
We formalize the notion of minimax regret in multiple steps.
First of all, $\Regret(x, \profile, \w)$ is the “regret” of selecting $x$ as a winner instead of choosing the optimal alternative under $\profile$ and $\w$:
\begin{align} 
	\Regret(x; \profile, \w) = \max_{y \in A} s(y; \profile,\w) - s(x; \profile, \w).
\end{align}
The pairwise maximum regret $\PMR(x,y;\pprofile,W)$ of $x$ relative to $y$ given the partial profile $\pprofile$ and the set of weights $W$
is the worst-case loss of choosing $x$ instead of $y$ under all possible realizations of the full profile {\em and} all possible instantiations of the weights:
\begin{align}
	\PMR(x,y; \pprofile, W) & = \max_{\w \in W} \max_{\profile \in C(\pprofile)} s(y; \profile,\w) - s(x; \profile,\w).
\end{align}

The maximum regret $\MaxR(x;\pprofile,W)$ is the worst-case loss of $x$. That is the loss occurred as the result of an adversarial selection of the complete profile $\profile \in C(\pprofile)$ and of the scoring vector $\w \in W$ that together maximize the loss between $x$ and the true winner under $\profile$ and $\w$:
\begin{align}
	\MaxR(x; \pprofile, W) & = \max_{y \in A} \PMR(x,y; \pprofile, W)\\
	& = \max_{\w \in W} \max_{\profile \in C(\profile)} \Regret(x; \profile, \w).
\end{align}

Finally,  $\MMR(\pprofile,W)$ is the value of minimax regret under $\pprofile$ and $W$, obtained when recommending a minimax optimal alternative $x^*_{\pprofile, W} \in A^*_{\pprofile, W}$:

\begin{align}
	\MMR(\pprofile,W) & = \min_{x \in A} \MaxR(x;\pprofile,W); \\
	A^*_{\pprofile, W} & = \argmin_{x \in A} \MaxR(x;\pprofile,W).
\end{align}
By picking as consensus choice
an alternative associated with minimax regret, we can provide a recommendation that gives worst-case guarantees, giving some robustness in face of uncertainty (due to both not knowing the agent preferences and the weights used in the aggregation). 
In cases of ties in minimax regret, we can either decide to return all minimax alternatives $A^*_{\pprofile, W}$ as winners or to pick just one of them using some tie-breaking strategy.

@@

Recall that, in the computation of $s(x; \profile, \w)$, $w_{v_j(x)}$ represents the score that $x$ obtains in the ranking $v_j$ (see \cref{eq:srule}).
Since scoring rules are additively decomposable, we can consider separately the contribution of each agent to the total score. Thus, we can write the actual regret of choosing $x$ instead of $y$ as
\[
s(y; \profile,\w) - s(x; \profile, \w) = \sum_{j=1}^n w_{v_j(y)} - w_{v_j(x)},
\]
and we can rewrite $\PMR$ as follows:
\begin{align}
	& \PMR(x,y; \pprofile, W) = \max_{\w \in W} \max_{\profile \in C(\pprofile)} [ s(y; \profile,\w) - s(x; \profile,\w) ] \\
	& =  \max_{\w \in W} \sum_{j=1}^{n} \max_{v_j \in C(\succ_j^p)} [w_{v_j(y)} - w_{v_j(x)}]. \\
\end{align}
Note that in general the inner max depends on the weights chosen by the outer max.

@@

The proposition can also be understood by observing that in the case $(x \pprefeq_j y)$, $\beta$ is the number of alternatives incomparable with both $x$ and $y$.
\begin{proposition}
	The $\PMR$ can be written as:
	\begin{align} 
		\PMR(x,y; \pprofile, W)  
		& = \max_{w \in W} \sum_{j=1}^n w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} \\ 
		& = \max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i. 
	\end{align}
	where $\hat{\alpha}_{r}^{y}$ (resp. $\hat{\alpha}_{r}^{x}$)  is the number of times $y$ (resp. $x$) is at rank $r$ in the complete profile $\hat{\profile}$ as determined in \cref{claim:rankPMR}. %\ref{claim:completion}.
\end{proposition}
The last proposition shows that PMR is linear in the weights.
Recall that the preferences of the chair are encoded with linear constraints $\Co_{W}$.
The pairwise max regret $\PMR(x,y; \pprofile,W)$ can be obtained as the solution of the following linear program defined on the variables $w_1, …, w_m$, which represent the weights attached to different positions:
\begin{align}
	\max_{\w} & \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r}\\
	\text{ s.t. } & \text{\cref{eq:monotone}} \text{ and } \text{\cref{eq:convexity}} \text{ and } \Co_W.
\end{align}
Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables 
(we leave $w_{1}$ and $w_{m}$ in the LP just for clarity of presentation).

@@

A bit more discussion is needed about {\em questions asked to the chair}.
Such questions aim at refining our knowledge about the scoring rule; a response gives us a constraint on the weight vector $\w$.
In particular, we want to obtain constraints of the type
\begin{align}
	w_{r} - w_{r+1} \geq \lambda (w_{r+1} - w_{r+2})
\end{align}
for $r \in \{1,\ldots,m-2\}$, relating the difference between the importance of ranks $r$ and $r+1$ with the difference between ranks $r+1$ and $r+2$.

>>>>

Let $V$ be the set of possible preference profiles (the union, for any integer $n$, of the $n$-fold Cartesian product of the linear orders over the alternatives).
A social choice function $f : V \rightarrow \powersetz{A}$ associates a profile with a set of winners, where $\powersetz{A}$ represents the set of subsets of $A$ except for the empty set (sets are used for tied winners).
Among the many possible social choice functions, we consider {\em positional scoring rules (PSR)}, which attach weights to positions according to the vector $(w_1, \ldots, w_m)$, also known as the {\em scoring vector}.
Let $\alpha^{x}_r$ be the number of times that alternative $x$ was ranked in the $r$-th position.
An alternative obtains a score that depends on the rank obtained in each of the preference orders:
\begin{align}
	\label{eq:srule}
	s(x; \profile, \w) = \sum_{j=1}^{n} w_{v_j(x)}
	= \sum_{r=1}^{m} \alpha^{x}_r w_r 
\end{align}
The winners are the alternatives with the highest score.

In this work we assume fixed, but unknown to us, a profile $\profile^*$ representing the preferences $\pref_j^*$ of the agents, and a weight vector $\w^*$ representing the preferences of the chair. We will reason about partial preference information concerning those objects.
At a given time, our knowledge of agent $j$'s preference is encoded by a partial order over the alternatives, thus a transitive and asymmetric binary relation, denoted by $\ppref_j$. 
In this work we assume that preference information is truthful, i.e. $a \ppref_j b ⇒ a \pref_j^* b$.
%We use $\prefinc$ to denote incomparability, that is $a \prefinc_{j} b$ iff $a \nppref_j b \wedge b \nppref_j a$.
An incomplete profile $\pprofile = (\ppref_1, \ldots, \ppref_n)$ maps each agent to a partial preference.

@@

We also assume that the weights of the scoring rule are only partially specified.
Therefore, the vector $(w_1,\ldots,w_m)$  is not known but we are given a set of constraints restraining the possible values that weights can take.
We consider a decreasing sequence of weights $1 = w_1 ≥ w_2 ≥ … ≥ w_m = 0$, a natural assumption, as it is better to be ranked first than second, second than third, etc.

@@

\section{Robust winner determination}
\label{sec:mmr}
In this paper, we consider that both the agent preferences and the preferences of the chair about the voting rule are incomplete.
As reviewed in \cref{sec:related}, some authors have considered possible and necessary winners assuming 
either a partial profile or assuming an incompletely specified scoring rule.

Notice that often there are no necessary winners and too many possible winners; it is therefore useful to declare a winner given partial information.
As a decision criterion to determine a winner, we propose to use minimax regret \citep{Savage1954}. 
This decision criterion has been used for robust optimization under data uncertainty \citep{Kouvelis1997} as well as in decision-making with uncertain utility values \citep{Salo2001,Boutilier2006}.
In particular, \citet{Lu2011} have adopted minimax regret for winner determination in social choice where
the preferences of agents that are only partially known, while the social choice function is predetermined and known.

@@

The maximum regret is considered by assuming that an adversary can both 1) extend the partial profile $\pprofile$ into a complete profile, and 2) instantiate the weights choosing among any weight vector in $\pw$, where $\pprofile$ and $\pw$ represent our knowledge so far.
We formalize the notion of minimax regret in multiple steps.
First of all, $\Regret(x, \profile, \w)$ is the “regret” of selecting $x$ as a winner instead of choosing the optimal alternative under $\profile$ and $\w$: $\Regret(x; \profile, \w) = \max_{y \in A} s(y; \profile,\w) - s(x; \profile, \w)$.
The pairwise maximum regret $\PMR(x,y;\pprofile,W)$ of $x$ relative to $y$ given the partial profile $\pprofile$ and the set of weights $W$
is the worst-case loss of choosing $x$ instead of $y$ under all possible realizations of the full profile {\em and} all possible instantiations of the weights: $\PMR(x,y; \pprofile, W) & = \max_{\w \in W} \max_{\profile \in C(\pprofile)} s(y; \profile,\w) - s(x; \profile,\w)$.
The maximum regret $\MaxR(x;\pprofile,W)$ is the worst-case loss of $x$. That is the loss occurred as the result of an adversarial selection of the complete profile $\profile \in C(\pprofile)$ and of the scoring vector $\w \in W$ that together maximize the loss between $x$ and the true winner under $\profile$ and $\w$: $\MaxR(x; \pprofile, W) = \max_{y \in A} \PMR(x,y; \pprofile, W) = \max_{\w \in W} \max_{\profile \in C(\profile)} \Regret(x; \profile, \w)$.
Finally,  $\MMR(\pprofile,W)$ is the value of minimax regret under $\pprofile$ and $W$, obtained when recommending a minimax optimal alternative $x^*_{\pprofile, W} \in A^*_{\pprofile, W}$: $\MMR(\pprofile,W) & = \min_{x \in A} \MaxR(x;\pprofile,W)$ and $A^*_{\pprofile, W} & = \argmin_{x \in A} \MaxR(x;\pprofile,W)$.
By picking as consensus choice
an alternative associated with minimax regret, we can provide a recommendation that gives worst-case guarantees, giving some robustness in face of uncertainty (due to both not knowing the agent preferences and the weights used in the aggregation). 
In cases of ties in minimax regret, we can either decide to return all minimax alternatives $A^*_{\pprofile, W}$ as winners or to pick just one of them using some tie-breaking strategy.

@@

Recall that, in the computation of $s(x; \profile, \w)$, $w_{v_j(x)}$ represents the score that $x$ obtains in the ranking $v_j$ (see \cref{eq:srule}).
Since scoring rules are additively decomposable, we can consider separately the contribution of each agent to the total score. Thus, we can write the actual regret of choosing $x$ instead of $y$ as $s(y; \profile,\w) - s(x; \profile, \w) = \sum_{j=1}^n w_{v_j(y)} - w_{v_j(x)}$, and we can rewrite $\PMR$ as $\PMR(x,y; \pprofile, W) = \max_{\w \in W} \max_{\profile \in C(\pprofile)} [ s(y; \profile,\w) - s(x; \profile,\w) ] =  \max_{\w \in W} \sum_{j=1}^{n} \max_{v_j \in C(\succ_j^p)} [w_{v_j(y)} - w_{v_j(x)}]$.
Note that in general the inner max depends on the weights chosen by the outer max.

@@

The proposition can also be understood by observing that in the case $(x \pprefeq_j y)$, $\beta$ is the number of alternatives incomparable with both $x$ and $y$.
\begin{proposition}
	The $\PMR$ can be written as:
	\begin{align} 
		\PMR(x,y; \pprofile, W)  
		& = \max_{w \in W} \sum_{j=1}^n w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} \\ 
		& = \max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i. 
	\end{align}
	where $\hat{\alpha}_{r}^{y}$ (resp. $\hat{\alpha}_{r}^{x}$)  is the number of times $y$ (resp. $x$) is at rank $r$ in the complete profile $\hat{\profile}$ as determined in \cref{claim:rankPMR}. %\ref{claim:completion}.
\end{proposition}
The last proposition shows that PMR is linear in the weights.
Recall that the preferences of the chair are encoded with linear constraints $\Co_{W}$.
The pairwise max regret $\PMR(x,y; \pprofile,W)$ can be obtained as the solution of the following linear program defined on the variables $w_1, …, w_m$, which represent the weights attached to different positions:
\begin{align}
	\max_{\w} & \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r}\\
	\text{ s.t. } & w_1 = 1, w_m = 0, \text{\cref{eq:convexity}} \text{ and } \Co_W.
\end{align}
%Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables 
%(we leave $w_{1}$ and $w_{m}$ in the LP just for clarity of presentation).

@@

A bit more discussion is needed about {\em questions asked to the chair}.
Such questions aim at refining our knowledge about the scoring rule; a response gives us a constraint on the weight vector $\w$.
In particular, we want to obtain constraints of the type $w_{r} - w_{r+1} \geq \lambda (w_{r+1} - w_{r+2})$ for $r \in \{1,\ldots,m-2\}$, relating the difference between the importance of ranks $r$ and $r+1$ with the difference between ranks $r+1$ and $r+2$.
